---
title: "Peer-graded Assignment: Course Project 1"
output: html_document
---

## R Markdown

### Set global option
```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
Sys.setlocale("LC_TIME","us")
```

### library
```{r,message=FALSE,warning=FALSE}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
```


### Loading and preprocessing the data
1. Load the data
```{r}
dat <- read.csv("activity.csv") 
```

2. transform the data class into "tbl_df"
```{r}
dat <- dplyr::as_data_frame(dat)
print(dat)
```

### What is mean total number of steps taken per day?
1. Calculate the total number of steps taken per day
```{r}
dat2 <- dat %>%
        dplyr::group_by(date) %>%
        dplyr::summarise(total_steps = sum(steps,na.rm=TRUE))
print(dat2)
```

2. Make a histogram of the total number of steps taken each day
```{r,fig.height=4}
g <- ggplot(dat2,aes(x=total_steps))
g + 
 geom_histogram(binwidth = 5000 ,fill="white",colour="black",boundary=0) + 
 labs(x="total steps",title="histogram of the total number of steps taken each day") +
 geom_vline(aes(xintercept=mean(dat2$total_steps),color="mean"),show.legend = TRUE) + 
 geom_vline(aes(xintercept=median(dat2$total_steps),color="median"),show.legend = TRUE) +
 scale_color_manual(name = "statistics", values = c(mean = "red",median = "blue"))
```

3. Calculate and report the mean and median of the total number of steps taken per day
```{r,comment=NULL}
mean <- mean(dat2$total_steps)
median <- median(dat2$total_steps)
c(mean=mean,median=median)
```

### What is the average daily activity pattern?

1. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r,fig.height=4}
dat3 <- dat %>%
        dplyr::group_by(interval) %>%
        dplyr::summarise(average_steps = mean(steps,na.rm=TRUE))
g <- ggplot(dat3,aes(x=interval,y=average_steps))
g + 
 geom_line(size = 1, colour="blue") + 
 labs(x="5-minute interval",y="average number of steps",
      title="the average daily activity pattern")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r,comment=NULL}
dat3 %>%
  dplyr::filter(average_steps == max(average_steps)) %>%
  as.data.frame
```

### Imputing missing values

1. Calculate and report the total number of missing values in the dataset
```{r,comment=NULL}
sum(is.na(dat))
```

2. Devise a strategy for filling in all of the missing values in the dataset. 

**Filling in all of the missing values by the average number of steps taken each day and each interval**

First, calculate the average number of steps taken each day
```{r}
ave_par_day <- dat %>%
        dplyr::group_by(date) %>%
        dplyr::summarise(average_day = mean(steps,na.rm=TRUE)) 

#If average valu is "NaN", set the mean of the total number of steps across all day
ave_all_day <- mean(ave_par_day$average_day,na.rm=TRUE)
ave_par_day <- ave_par_day %>%
  dplyr::mutate(average_day = ifelse(is.na(average_day),ave_all_day,average_day)) %>%
  dplyr::mutate(weight = average_day/mean(average_day))
ave_par_day
```

Second,  calculate the average number of steps taken each interval
```{r}
ave_par_interval <- dat %>%
        dplyr::group_by(interval) %>%
        dplyr::summarise(average_interval = mean(steps,na.rm=TRUE)) 
ave_par_interval
```

Next, joining the average number of steps taken each day and each interval,and calculate the interpolation
```{r}
interpolate_dat <- dat %>%
                   dplyr::inner_join(ave_par_day,by="date") %>% 
                   dplyr::inner_join(ave_par_interval,by="interval") %>% 
                   dplyr::mutate(fill = weight * average_interval) 
interpolate_dat
```

Finally, filling  in all of the missing values
```{r}
interpolate_dat <- interpolate_dat %>%        
   dplyr::mutate(interpolation = ifelse(is.na(steps),fill,steps)) %>%
   dplyr::select(-(average_day:fill))
interpolate_dat
```

3.Create a new dataset that is equal to the original dataset but with the missing data filled in
```{r}
interpolate_dat <- interpolate_dat %>%        
   dplyr::mutate(steps = interpolation) %>%
   dplyr::select(steps:interval)
interpolate_dat
```

4.Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r,fig.height=4}
interpolate_dat2 <- interpolate_dat %>%
        dplyr::group_by(date) %>%
        dplyr::summarise(total_steps = sum(steps,na.rm=TRUE))

g <- ggplot(interpolate_dat2,aes(x=total_steps))
g + 
 geom_histogram(binwidth = 5000 ,fill="white",colour="black",boundary=0) + 
 labs(x="total steps",
      title="histogram of the total number of steps with the missing data filled in ") +
 geom_vline(aes(xintercept=mean(dat2$total_steps),color="mean_NA"),
            linetype='dashed',show.legend = TRUE) +
 geom_vline(aes(xintercept=mean(interpolate_dat2$total_steps),color="mean"),
            size=1,show.legend = TRUE) + 
 geom_vline(aes(xintercept=median(dat2$total_steps),color="median_NA"),
            linetype='dashed',show.legend = TRUE) +
 geom_vline(aes(xintercept=median(interpolate_dat2$total_steps),color="median"),
            size=1,show.legend = TRUE) +
 scale_color_manual(name = "statistics",
            values = c(mean_NA = "red",mean = "green",median_NA = "blue",median = "green"))

```


Do these values differ from the estimates from the first part of the assignment?What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r}
mean_Fir <- mean(dat2$total_steps)
median_Fir <- median(dat2$total_steps)
mean_Sec <- mean(interpolate_dat2$total_steps)
median_Sec <- median(interpolate_dat2$total_steps)

df <- data.frame(mean=c(mean_Fir,mean_Sec,mean_Sec-mean_Fir),
                 median=c(median_Fir,median_Sec,median_Sec-median_Fir))
rownames(df) <- c("with the missing data","with the missing data filled in","difference")
df
```


### Are there differences in activity patterns between weekdays and weekends?

1.Create a new factor variable in the dataset with two levels – “weekday” and “weekend” 
```{r}
dat4 <- interpolate_dat %>%
  mutate(weekday = weekdays(lubridate::ymd(date),abbreviate=TRUE)) %>%
  mutate(weekend = factor(weekday,
   levels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"),
   labels = c("weekday", "weekday", "weekday", "weekday", "weekday", "weekend", "weekend")))
dat4
```

2. Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r,fig.height=4}
dat5 <- dat4 %>%
        dplyr::group_by(interval,weekend) %>%
        dplyr::summarise(average_steps = mean(steps,na.rm=TRUE))
g <- ggplot(dat5,aes(x=interval,y=average_steps,group=weekend))
g + 
 facet_grid(aes(group=weekend)) +
 geom_line(size = 1) + 
 labs(x="5-minute interval",y="average number of steps",
      title="the average daily activity pattern")
```